### 2016-01-25 inputs/results.tex
nelson action points
aenderungen-19:55

\subsection{Genotypes}
\label{sec:genotypes}  
GBS detected 32\,706, 40\,734 and 166\,418 SNP loci in Alfalfa-PV, Alfalfa-Med and Rice, respectively (Table~\ref{tab:descriptive_statistics}). The overall missing rate was 0.666, 0.596 and 0.534 in the three datasets. Supplementary Figure S1 shows the density distribution of missing rate per markers and per individuals in the complete datasets. The amount of missing marker genotypes varied with the threshold of maximum per-marker missing-rate allowed in the data (10\%, 20\%, 40\% and 70\%) and the proportion of artificial missing genotypes that were introduced (1\%, 5\%, 10\% and 20\%). Missing data point counts ranged from a minimum of 3\,878 in Alfalfa-PV with 10\% allowed and 1\% artificial missing genotypes, to a maximum of 5\,580\,616 in Rice with 70\% allowed and 20\% artificial missing genotypes. Over all datasets and allowed/artificial missing genotypes thresholds, the amount of missing data points to be imputed averaged 249\,405.\\
Minor Allele Frequency (MAF) was 0.172, 0.170 and 0.140 in Alfalfa-PV, Alfalfa-Med and Rice, respectively. The three datasets were therefore unbalanced with respect to genotype classes: Rice had the lowest MAF, while Alfalfa-PV and Alfalfa-Med had the smallest minor classes (minor homozygotes 3.5\% and 3.2\% of the total genotypes). Figure~\ref{fig:genotype_classes} reports the proportion of genotypes in each class in the complete datasets (100\%) and as a function of the maximum allowed missing rate per marker. The relative proportion of genotype classes remained approximately stable in rice. In contrast, the heterozygous class in alfalfa tended to become smaller with increasing proportion of allowed genotypes. This reflected the different GBS calling criteria for homozygous and heterozygous SNP: heterozygous SNPs required fewer overlapping reads to be called, while the implemented quality filters on alfalfa required a larger number of reads to call a homozygous loci. The average missing-rate and MAF in the 12 rice chromosomes were comparable to the whole-rice dataset.

@@

\subsection{Computation time}
\label{sec:computation_time}  
We measured the efficiency of the different imputation algorithms as the amount of time required to complete the imputation process. Only the implementation of RFI could leverage a multi-core/multi-thread environment, so that RFI computation times should be evaluated considering 10 CPUs used in parallel, while all other algorithms used only one CPU at a time.\\
Imputation efficiency (Figure~\ref{fig:time_analysis}) was assessed with respect to the gross dimension of the data set (i.e., number of markers $\times$ number of samples) and to the total number of missing genotypes to be imputed (e.g., 341\,097 missing genotypes for the Alfalfa-Med population with 40\% missing rate and 20\% artificial missing).
RFI required by far the longest computation times (in spite of 10-core/thread parallelization), which grew approximately exponentially ($O(e^N)$) with problem size, and easily required tens of hours for individual rice chromosomes, and hundred of hours (up to 937 hours) for the complete rice dataset. The second slowest algorithm was KNNI, with computation times growing approximately quadratically ($O(N^2)$) with problem complexity (the chosen KNNI implementation contains no heuristic method to prune the all vs. all distance calculation). KNNI was however significantly faster than RFI (on average 16.7 times faster), requiring a time ranging from 20.68 seconds (rice chromosome 7, 10\% allowed, 1\% artificial) to about 28 hours (rice complete dataset, 70\% allowed, 20\% artificial) to complete the imputation task.\\
All other algorithms were faster, with computation times growing linearly or logarithmically with problem size. Beagle and SVDI resulted in similar execution times. MNI was by far the fastest imputation algorithm, being scarcely affected by the size of the problem (computation times ranging from 0.23 to 31.29 seconds).

>>>>

\subsection{Genotypes}
\label{sec:genotypes}  
GBS detected 32\,706, 40\,734 and 166\,418 SNP loci in Alfalfa-PV, Alfalfa-Med and Rice, respectively (Table~\ref{tab:descriptive_statistics}). The overall missing rate was 0.666, 0.596 and 0.534 in the three datasets. Supplementary Figure S1 shows the density distribution of missing rate per markers in the complete datasets. The amount of missing marker genotypes varied with the threshold of maximum per-marker missing-rate allowed in the data (10\%, 20\%, 40\% and 70\%) and the proportion of artificial missing genotypes that were introduced (1\%, 5\%, 10\% and 20\%). Missing data point counts ranged from a minimum of 3\,878 in Alfalfa-PV with 10\% allowed and 1\% artificial missing genotypes, to a maximum of 5\,580\,616 in Rice with 70\% allowed and 20\% artificial missing genotypes. Over all datasets and allowed/artificial missing genotypes thresholds, the amount of missing data points to be imputed averaged 249\,405.\\
Minor Allele Frequency (MAF) was 0.172, 0.170 and 0.140 in Alfalfa-PV, Alfalfa-Med and Rice, respectively. The three datasets were therefore unbalanced with respect to genotype classes: Rice had the lowest MAF, while Alfalfa-PV and Alfalfa-Med had the smallest minor classes (minor homozygotes 3.5\% and 3.2\% of the total genotypes). Figure~\ref{fig:genotype_classes} reports the proportion of genotypes in each class in the complete datasets (100\%) and as a function of the maximum allowed missing rate per marker. The relative proportion of genotype classes remained approximately stable in rice. In contrast, the heterozygous class in alfalfa tended to become smaller with increasing proportion of allowed genotypes. This reflected the different GBS calling criteria for homozygous and heterozygous SNP: heterozygous SNPs required fewer overlapping reads to be called, while the implemented quality filters on alfalfa required a larger number of reads to call a homozygous loci. The average missing-rate and MAF in the 12 rice chromosomes were comparable to the whole-rice dataset.

@@

\subsection{Computation time}
\label{sec:computation_time}  
The amount of time required to complete the imputation process was recorded for each method. Only the implementation of RFI could leverage a multi-core/multi-thread environment, so that RFI computation times should be evaluated considering 10 CPUs used in parallel, while all other algorithms used only one CPU at a time.\\
Imputation efficiency (Figure~\ref{fig:time_analysis}) was assessed with respect to the gross dimension of the data set (i.e., number of markers $\times$ number of samples) and to the total number of missing genotypes to be imputed (e.g., 341\,097 missing genotypes for the Alfalfa-Med population with 40\% missing rate and 20\% artificial missing).
RFI required by far the longest computation times (in spite of 10-core/thread parallelization), which grew approximately exponentially ($O(e^N)$) with problem size, and easily required tens of hours for individual rice chromosomes, and hundred of hours (up to 937 hours) for the complete rice dataset. The second slowest algorithm was KNNI, with computation times growing approximately quadratically ($O(N^2)$) with problem complexity (the chosen KNNI implementation contains no heuristic method to prune the all vs. all distance calculation). KNNI was however significantly faster than RFI (on average 16.7 times faster), requiring a time ranging from 20.68 seconds (rice chromosome 7, 10\% allowed, 1\% artificial) to about 28 hours (rice complete dataset, 70\% allowed, 20\% artificial) to complete the imputation task.\\
All other algorithms were faster, with computation times growing linearly or logarithmically with problem size. Beagle and SVDI resulted in similar execution times. MNI was by far the fastest imputation algorithm, being scarcely affected by the size of the problem (computation times ranging from 0.23 to 31.29 seconds).
